
\documentclass[aspectratio=169,professionalfonts]{beamer}
\usetheme{default}

% Basic packages
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath, amssymb}
\usepackage{array}
\usepackage{listings}
\usepackage{csquotes}
\usepackage{xspace}

% Beamer metadata
\title{Getting Research Done with ChatGPT Plus and Modern AI}
\subtitle{A one hour training for quantitative social sciences}
\author{Juan Pablo Posada Aparicio}
\institute{Institute for Replication, University of Ottawa}
\date{October 2025}

% Utilities
\newcommand{\feature}[1]{\textbf{#1}}
\newcommand{\demo}{\textbf{Demo. }}
\newcommand{\notepara}[1]{\small\textit{#1}}

\begin{document}

%-----------------------
\begin{frame}
  \titlepage
\end{frame}

%-----------------------
\begin{frame}{Session goals}
  \begin{itemize}
    \item Understand what large language models are, at a practical level.
    \item Learn ChatGPT Plus capabilities that matter for research.
    \item See how to plug AI into literature, coding, data, writing, and review.
    \item Get copy paste demo prompts you can run today.
    \item Leave with a short checklist to use AI responsibly and reproducibly.
  \end{itemize}
\end{frame}

%-----------------------
\begin{frame}{Roadmap, about 60 minutes}
  \begin{itemize}
    \item 5 minutes. LLMs in plain language and prompting basics.
    \item 10 minutes. What is in ChatGPT Plus today.
    \item 30 minutes. Research workflow playbook, literature to code to writing.
    \item 10 minutes. Deep Research and Agent Mode.
    \item 5 minutes. Guardrails, reproducibility, and Q and A.
    \item 5 minutes. Codex, quick version, at the very end.
  \end{itemize}
\end{frame}

%-----------------------
\section{LLMs in plain language}

\begin{frame}{What is a large language model}
  \begin{itemize}
    \item A model that predicts the next token in context. It turns that simple mechanism into useful tasks by following instructions.
    \item Strengths. Language understanding, code generation, pattern completion, transformation.
    \item Weaknesses. Can be wrong or overconfident. May miss recent facts without browsing or tools. Needs clear instructions and constraints.
    \item Your edge. Provide context, constraints, and examples. Iterate and verify.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Prompting that actually works}
  \begin{itemize}
    \item Set role and objective. Then give the smallest possible working example.
    \item Ask for structured outputs you can reuse. For example markdown, csv, json, or a full script.
    \item Use iterate and verify. Ask for one improvement at a time.
  \end{itemize}
\end{frame}

%-----------------------
\section{What is in ChatGPT Plus}

\begin{frame}{ChatGPT Plus, the useful bits}
  \begin{itemize}
    \item \feature{Advanced Data Analysis}. Write and run Python inside the chat. Upload files. Produce figures and tables.
    \item \feature{Browsing}. Read the web for current facts when you ask for it.
    \item \feature{Deep Research}. Multi step web research that collects and organizes findings before answering.
    \item \feature{Custom GPTs}. Create specialized assistants with your instructions and files.
    \item \feature{Agent Mode}. Let ChatGPT plan and execute multi step tasks with tools, while you supervise.
  \end{itemize}
\end{frame}

%-----------------------
\section{Research workflow playbook}

\begin{frame}{Literature discovery and synthesis}
  \begin{itemize}
    \item Ask for a scaffold first. Key theories, main debates, measurement, common datasets.
    \item Then ask for concrete items. Seminal papers, recent systematic reviews, and data portals.
    \item Use browsing or Deep Research when you need current sources.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, literature}
\demo Copy and run one prompt.
\begin{verbatim}
You are my research assistant. Topic: the effect of community radios on
forced displacement in Colombia. Output a literature scaffold with
sections: theory, identification strategies seen in the field,
measurement choices, typical datasets, open questions. Limit to 10 bullets.
\end{verbatim}

\demo With browsing or Deep Research on.
\begin{verbatim}
Find 8 recent and credible sources on community radios and violence.
Return a table with columns: citation in APA, year, identification
strategy, sample, main finding, link. Put uncertainties in a final note.
\end{verbatim}
\end{frame}

\begin{frame}{Data access and documentation}
  \begin{itemize}
    \item Ask for a data dictionary plan. Variable names, units, refresh cadence, and owners.
    \item Have the model draft a README template for each dataset.
    \item Convert messy documentation into a clean table you can ship.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, data documentation}
\demo Dictionary and README.
\begin{verbatim}
Create a minimal data dictionary schema for a panel of municipalities by
month. Include: name, label, type, units, allowed values, missing policy,
owner, refresh cadence, source URL.
\end{verbatim}

\demo Turn messy text into a table.
\begin{verbatim}
Extract a clean csv with columns variable, definition, unit, period
from the following pasted notes. If needed, ask one clarifying question.
\end{verbatim}
\end{frame}

\begin{frame}{Cleaning and analysis with Advanced Data Analysis}
  \begin{itemize}
    \item Upload a small sample first. Ask for a quick EDA, missingness, and obvious anomalies.
    \item Ask for clean, reproducible code with comments and a final function.
    \item Iterate to reach publication grade figures.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, cleaning and EDA}
\demo Quick EDA on upload.
\begin{verbatim}
Run a 10 minute EDA on the uploaded csv. Show schema, missingness,
basic summaries, and 3 plots that reveal structure. Save code to a .py
file I can download.
\end{verbatim}

\demo Robust cleaning function.
\begin{verbatim}
Write a Python function clean_df(df) that trims whitespace, parses dates,
standardizes municipality names using a provided mapping, and validates
unique keys (muni_id, month). Return cleaned df and a log of fixes.
\end{verbatim}
\end{frame}

%-----------------------
\section{Interpreting results and writing}

\begin{frame}{Interpreting results and writing}
  \begin{itemize}
    \item Paste tables or model summaries and ask for a plain English interpretation.
    \item Ask for a short abstract or a paragraph that highlights identification and magnitude.
    \item Use a style checklist. Clarity, claims tied to evidence, and limits.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, interpretation and writing}
\demo Interpret a regression.
\begin{verbatim}
Given this regression table, explain the sign, magnitude, and precision
of key coefficients. Avoid causal language unless identification is explicit.
\end{verbatim}

\demo Draft a concise abstract.
\begin{verbatim}
Draft a 150 word abstract given these bullet points. Emphasize design,
data, and main effect with uncertainty. Return a single paragraph.
\end{verbatim}
\end{frame}

%-----------------------
\section{Peer review and reproducibility}

\begin{frame}{Peer review and reproducibility}
  \begin{itemize}
    \item Ask the model to act as a critical but fair reviewer.
    \item Have it produce a replication log template, aligned with your PAP.
    \item Use it to standardize robustness checks and pre analysis notes.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, review and logs}
\demo Act as a reviewer.
\begin{verbatim}
Act as a critical but fair referee for the following methods section.
List 5 specific concerns and 3 concrete suggestions to improve validity.
\end{verbatim}

\demo Replication log template.
\begin{verbatim}
Create a csv template for a replication log with columns for coding
errors, computational reproducibility, two robustness checks, and notes.
Keep column names short and consistent.
\end{verbatim}
\end{frame}

%-----------------------
\section{Advanced features}

\begin{frame}{Deep Research, when to use it}
  \begin{itemize}
    \item Use when you want structured web research with sources collected first.
    \item Ask for constraints. Scope, credibility criteria, and deliverable format.
    \item Expect longer runs. Use it for scoping reviews or policy scans.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, Deep Research}
\demo Scoping review pattern.
\begin{verbatim}
Use Deep Research to scan policy reports and peer reviewed papers on
municipal radio and governance outcomes in Latin America after 2010.
Return a matrix of findings by country and outcome with links.
\end{verbatim}
\end{frame}

\begin{frame}{Agent Mode, supervised autonomy}
  \begin{itemize}
    \item Plan and execute multi step tasks across tools, under your supervision.
    \item Good for browse, download data, analyze, and assemble a draft.
    \item Start with small, auditable tasks. Keep humans in the loop.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Demo prompts, Agent Mode}
\demo End to end micro task.
\begin{verbatim}
Plan and run a short workflow. Find two official data sources on radio
coverage by municipality in Colombia, download small samples, summarize
coverage over time for Antioquia, and draft a 5 bullet summary with links.
Ask before any login or large downloads.
\end{verbatim}
\end{frame}

%-----------------------
\section{Guardrails}

\begin{frame}{Responsible and reproducible use}
  \begin{itemize}
    \item Track provenance. Record what was prompted, what data was used, and when.
    \item Never paste sensitive data unless policy allows it. Prefer local tools for secrets.
    \item Verify everything that affects results or claims.
    \item Save outputs in plain text files. Scripts, logs, and markdown notes.
    \item Be clear about where AI helped, in acknowledgments or methods.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{A tiny checklist}
\begin{verbatim}
1. Did I provide sufficient context and constraints.
2. Did I verify claims and numbers against sources.
3. Are scripts and logs saved with dates and versions.
4. Can a colleague rerun this without me.
5. Is sensitive data handled according to policy.
\end{verbatim}
\end{frame}

%-----------------------
\section{Appendix, prompt bank}

\begin{frame}[fragile]{Prompt bank, quick copy}
\begin{verbatim}
Rewrite this paragraph for clarity and precision. Keep technical terms.
Return two variants and a diff style list of changes.
\end{verbatim}

\begin{verbatim}
You are a code reviewer. Read this Stata do file and list 5 risks.
Then propose 3 refactors that reduce runtime or errors.
\end{verbatim}

\begin{verbatim}
Given this figure and caption, propose one alternative visualization
and write the exact matplotlib or ggplot code.
\end{verbatim}
\end{frame}

%-----------------------
\section{Codex, quick version}

\begin{frame}{Codex, what it is and when to use it}
  \begin{itemize}
    \item Codex is OpenAI's coding agent for working with your repos.
    \item Use chat for questions and small snippets. Use Codex when you want edits in files and simple automations.
    \item Keep your repo under git. Review diffs. You stay in control.
  \end{itemize}
\end{frame}

\begin{frame}[fragile]{Codex CLI, install and start}
\begin{verbatim}
# macOS or Linux
npm install -g @openai/codex
# or
brew install codex

# run
codex
\end{verbatim}
\notepara{On Windows, prefer WSL.}
\end{frame}

\begin{frame}[fragile]{Two demos you can run}
\demo Summarize a csv.
\begin{verbatim}
Read data.csv here. Report column types, missingness,
and five key stats per numeric column. Save a markdown summary.
\end{verbatim}

\demo Explain code you did not write.
\begin{verbatim}
Explain what analysis.py does, step by step, and list two risks.
\end{verbatim}
\end{frame}

\begin{frame}[fragile]{Tiny command cheat sheet}
\begin{verbatim}
codex "explain this codebase"
codex exec "fix the failing test"
/model gpt-5-codex
/approvals
\end{verbatim}
\end{frame}

%-----------------------
\begin{frame}{Q and A}
  \centering Thank you.
\end{frame}

\end{document}
